---
title: "A3: SPL Library Checkouts"
author: "Jorey Papritz"
date: "03/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

### Introduction

  The data I am choosing to analyze is checkout data that includes books checked out at least 10 times a month from the Seattle Public Library from 2017 to 2023.I am interested in looking at the overall trends in checkout patterns for physical and digital materials, as well as the most popular material types and the top five most checked out books. By analyzing these trends, I hope to gain insights into the reading habits and preferences of people using the library, as well as any notable changes or patterns over time. Analyzing the data could also reveal any potential biases or inequalities in the collection and usage of library materials, which is an important ethical consideration to keep in mind.

### Summary Information

  The average number of digital checkouts was 25.4 which was slightly higher than physical checkouts. The year with the highest number of ebook checkouts was 2022 with 1962294 checkouts. 
  The number of print/digital book checkouts varied over time. The total checkouts of print and digital materials increased from 2017 to 2019, but decreased from 2019 to 2022.
  The most popular material type was 'BOOK' with a total of 2587631 checkouts.
  The five most checked-out items and their checkout counts varied over time. For example, in 2017, the most checked-out item was 'The Girl on the Train' with 3298 checkouts, while in 2022, it was 'Where the Crawdads Sing' with 1772 checkouts.

### The Dataset

The collector and publisher of the data is the Seattle Public Library. This dataset contains physical and digital book borrowing records from 2017 to the present. It contains information such as year borrowed, month borrowed, date borrowed, title, author, ISBN, year of publication, publisher and type of material (book, audiobook, DVD, e-book, etc.). The dataset contains items checked out at least 10 times a month from 2017 to 2023. This data has 12 rows and 816354 columns. This data are collected electronically by the Seattle Public Library. When a book is loaned out, the system records information such as the title, author, year of publication, and other details about the book. The data was collected to gain insight into the circulation patterns and interests of Seattle Public Library customers. It can be used to make decisions about purchasing new materials, managing collections, and planning library programs and services. This record contains identifiable information, such as name, address, and other user data. Any research using this data must ensure the privacy and security of this information. Any publicly released data must also be made anonymous to protect individual privacy. A limitation of this dataset is that it only reflects the circulation patterns of Seattle Public Library users. It does not necessarily represent the borrowing behavior of library users in other parts of the US. Additionally, the dataset might target a specific type of book or a specific customer segment. For example, if a library has a large collection of science fiction books, those books are likely to be checked out more often than other types of books. Also, this dataset does not contain any information about why users view books. Some books may be borrowed for academic or research purposes rather than for leisure reading.


### First Trends Over Time Chart

The chart displayed shows the trends in physical and digital checkouts from the Seattle Public Library from 2017 up until 2023. By looking at the chart, it can be observed that physical book checkouts have been decreasing over time while digital book checkouts have been growing. This suggests a shift in the reading habits of library users towards digital materials. This is very likely due to the COVID-19 pandemic in 2020. This chart provides useful information about the overall trends in checkout patterns for physical and digital materials, and gives the Seattle Public Library an idea of how to adjust their collection and services in order to meet the changing needs of their users.

``````{r, echo = FALSE, warning = FALSE, messages = FALSE, code = readLines("Chart-1.R")}
``````

### Second Trends Over Time Chart

  The second graph is a line plot that visualizes the trends in checkout patterns for physical books, audiobooks, and ebooks from the Seattle Public Library from 2017 to 2023. The chart shows three different colored lines, red for physical books, green for audiobooks, and blue for ebooks, to display how the total checkouts for each material type have changed over time.
  Compared to the first graph, which only showed the trends for physical and digital materials, this graph shows a more detailed breakdown of the types of materials being checked out from the library. It helps us understand which specific types of materials are most popular among library users and how their usage has changed over time.
  From the graph, we can see that physical books have consistently been the most checked out material type since 2017, followed by ebooks and then audiobooks. However, the trends in usage for each material type have been different. While physical books have shown a declining trend, ebooks and audiobooks have shown increasing trends. This is due to the COVID Pandemic in 2020 that led to people reading books using digital methods as people weren't allowed to leave their houses to go to the library.
  This graph helps librarians have a better understanding of the reading habits/preferences of their users and how they can optimize their collection and services to meet these changing needs. It also reveals potential biases or inequalities in the collection and usage of library materials, which can help administrators address these issues to make their library more accessible to all users.

``````{r, echo = FALSE, warning = FALSE, messages = FALSE, code = readLines("Chart-2.R")}
```

### Your Choice

  The chart is a bar plot that displays the top 5 most popular material types in the Seattle Public Library based on the total number of checkouts from 2017 to 2023.
  The first two graphs show trends over time, while the third graph shows a snapshot of the most popular material types at any given time. The first two graphs are useful for tracking changes in reader preferences and usage patterns over time, while the third graph provides better insight into the most popular material types at a particular point in time.
  I included this chart because it provides valuable insights into the types of materials that are most in demand among library users. It also helps librarians make good decisions about how to allocate their resources and improve their collection to meet the needs and preferences of their users.
  From the chart, we can see that books are the most popular material type by far, with over 7 million checkouts in the period analyzed. This is followed by ebooks, and then audiobooks. DVDs and music CDs were the fourth and fifth most popular material types.
  The chart might highlight potential biases/inequalities in the collection library materials. For example, the fact that books are by far the most popular material type suggests that the library may need to invest more in promoting other types of materials to serve more people.

``````{r, echo = FALSE, warning = FALSE, code = readLines("Chart-3.R")}
```

